---
title: "Power Calc NLP Hypercap Ref Std"
format: html
editor: visual
---

## Power Calculation for the Reference Standard Annotation

MIMIC IV; NLP to classify NHAMCS RVC visits.

```{r}
#| label: chart-review-sample-calc-cell-001
#| message: false
#| warning: false
# Deterministic runtime contract: fail fast when required packages are absent.
required_pkgs <- c("presize", "kappaSize", "irr")
missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_pkgs) > 0) {
  stop(
    paste0(
      "Missing required R packages: ",
      paste(missing_pkgs, collapse = ", "),
      ". Install them in your managed R environment (e.g., install.packages(c(",
      paste(shQuote(missing_pkgs), collapse = ", "),
      "))) and rerender."
    )
  )
}
```

## Assumptions used in planning

- Target agreement: $\kappa \approx 0.7-0.8$
- Desired precision: 95% confidence interval half-width around $\pm 0.10$
- Multi-class label distribution approximated from pilot review (5-category simplification for planning)
- Resulting planning sample is approximately **n ≈ 160** encounters

1)  Precision via `presize`: target-width planning and expected CI width at **n = 60**

```{r}
#| label: chart-review-sample-calc-cell-002
library(presize)
planning_precision <- prec_kappa(
  kappa = 0.70,
  conf.width = 0.20,
  raters = 2,
  n_category = 5,
  props = c(0.5, 0.2, 0.1, 0.05, 0.15),
  conf.level = 0.95
)

n60_precision <- prec_kappa(
  kappa = 0.70,
  n = 60,
  raters = 2,
  n_category = 5,
  props = c(0.5, 0.2, 0.1, 0.05, 0.15),
  conf.level = 0.95
)

list(
  required_n_for_target_width = planning_precision,
  expected_ci_width_with_n60 = n60_precision
)
```

And

2)  Same via kappaSize (explicit CI bounds)

```{r}
#| label: chart-review-sample-calc-cell-003
library(kappaSize)
CI5Cats(kappa0 = 0.80, kappaL = 0.70, kappaU = 0.90,
        props = c(0.5, 0.2, 0.1, 0.1 , 0.1), raters = 2, alpha = 0.05)
```

and

3)  Hypothesis-based power (Flack et al. 1988):

```{r}
#| label: chart-review-sample-calc-cell-004
#    test H0: kappa = 0.50 vs H1: kappa = 0.70, 80% power, 5 categories
library(irr)
mrg <- round(c(0.5, 0.2, 0.1, 0.1, 0.1), 12)
mrg[length(mrg)] <- 1 - sum(mrg[-length(mrg)])
irr::N2.cohen.kappa(mrg, k1=0.70, k0=0.50, alpha=0.05, power=0.80)
```

and

4)  Expected 95% CI width if only **n = 60** encounters are annotated

```{r}
#| label: chart-review-sample-calc-cell-005
#| message: false
#| warning: false
set.seed(1234)

n_annotations <- 60L
n_sim <- 5000L
planning_props <- c(0.5, 0.2, 0.1, 0.05, 0.15)
kappa_targets <- c(0.70, 0.75, 0.80)

simulate_kappa_once <- function(n, kappa_target, props) {
  # Draw rater-1 labels from planning category distribution.
  r1 <- sample.int(length(props), size = n, replace = TRUE, prob = props)
  # Agreement-mixture model: under this construction expected kappa ~= kappa_target.
  agree <- rbinom(n, size = 1, prob = kappa_target)
  r2_independent <- sample.int(length(props), size = n, replace = TRUE, prob = props)
  r2 <- ifelse(agree == 1, r1, r2_independent)
  irr::kappa2(data.frame(r1 = r1, r2 = r2), weight = "unweighted")$value
}

simulate_kappa_distribution <- function(n, kappa_target, props, n_sim) {
  replicate(n_sim, simulate_kappa_once(n, kappa_target, props))
}

scenario_summary <- do.call(
  rbind,
  lapply(kappa_targets, function(kappa_target) {
    kappa_draws <- simulate_kappa_distribution(
      n = n_annotations,
      kappa_target = kappa_target,
      props = planning_props,
      n_sim = n_sim
    )
    ci <- quantile(kappa_draws, probs = c(0.025, 0.5, 0.975), na.rm = TRUE)
    data.frame(
      n_annotations = n_annotations,
      kappa_target = kappa_target,
      median_kappa = unname(ci[2]),
      ci_lower = unname(ci[1]),
      ci_upper = unname(ci[3]),
      ci_width = unname(ci[3] - ci[1]),
      ci_half_width = unname((ci[3] - ci[1]) / 2),
      na_rate = mean(is.na(kappa_draws))
    )
  })
)

scenario_summary[] <- lapply(scenario_summary, function(x) {
  if (is.numeric(x)) round(x, 3) else x
})
scenario_summary
```
